---
title: "Group_Project"
format: html
editor: visual
---

## Quarto

## Libraries:

```{r}
library(haven)
library(dplyr)
library(ggplot2)
```

Data:

```{r}
#| echo: false
data <- read_dta("ZA7575.dta")
```

```{r}
#| echo: false
summary(data)
head(data)


column_info <- data.frame(
  column_name = colnames(data),
  column_label = sapply(data, function(x) attr(x, "label"))
)

column_info

data <- data %>%
  mutate(country_name = case_when(
    q1_1  == 1 ~ "Belgium",
    q1_2  == 1 ~ "Denmark",
    q1_3  == 1 ~ "Germany",
    q1_4  == 1 ~ "Greece",
    q1_5  == 1 ~ "Spain",
    q1_6  == 1 ~ "France",
    q1_7  == 1 ~ "Ireland",
    q1_8  == 1 ~ "Italy",
    q1_9  == 1 ~ "Luxembourg",
    q1_10 == 1 ~ "Netherlands",
    q1_11 == 1 ~ "Portugal",
    q1_12 == 1 ~ "United Kingdom",
    q1_13 == 1 ~ "Austria",
    q1_14 == 1 ~ "Sweden",
    q1_15 == 1 ~ "Finland",
    q1_16 == 1 ~ "Cyprus",
    q1_17 == 1 ~ "Czech Republic",
    q1_18 == 1 ~ "Estonia",
    q1_19 == 1 ~ "Hungary",
    q1_20 == 1 ~ "Latvia",
    q1_21 == 1 ~ "Lithuania",
    q1_22 == 1 ~ "Malta",
    q1_23 == 1 ~ "Poland",
    q1_24 == 1 ~ "Slovakia",
    q1_25 == 1 ~ "Slovenia",
    q1_26 == 1 ~ "Bulgaria",
    q1_27 == 1 ~ "Romania",
    q1_28 == 1 ~ "Croatia",
    q1_29 == 1 ~ "Other countries",
    q1_30 == 1 ~ "DK",
    TRUE ~ NA_character_  # Falls keine der Bedingungen zutrifft
  ))




```

## legal framework

QC19:Do you think that transgender persons should be able to change their civil documents to match their inner gender identity?(1=yes, 2=no,3=dk) QC20: Do you believe that official documents, like passports and birth certificates, should have a third option, such as X or O (other) beside male (M) and female (F)for those persons who do not identify as female and male? (1=yes, 2=no,3=dk) QC15_1: to what extend to do you agree or disagree with the following statement? Gay, lesbian and bisexual people should have the same rights as heterosexual people (1 2 3 4 5) QC15_2: to what extend to do you agree or disagree with the following statement? here is nothing wrong in a sexual relationship between two persons of the same sex(1 2 3 4 5) QC15_3: to what extend to do you agree or disagree with the following statement? Same sex marriages should be allowed throughout europe (1 2 3 4 5) QC4_7: In (OUR COUNTRY) when a company wants to hire someone and has the choice between two candidates with equal skills and qualifications, which of the following criteria may, in your opinion, put one candidate at a disadvantage? The candidate’s gender or sex (man or woman) QC4_8: In (OUR COUNTRY) when a company wants to hire someone and has the choice between two candidates with equal skills and qualifications, which of the following criteria may, in your opinion, put one candidate at a disadvantage? The candidate’s gender identity (being transgender) QC5_3: in the last 12 month have you done: You have joined an association or campaign that defends people against discrimination? (1=yes, 2=no,3=dk) QC7: Using a scale from 1 to 10, please tell me if you think that the efforts made in (OUR COUNTRY) to fight all forms of discrimination are effective. ‘1’ means you consider that these efforts are "not at all effective", and ‘10’ that these efforts are "very effective".

```{r}
#| echo: false

legal_framework_data <- data %>%
  select(
    qc19,        # TRANSGENDER - CHANGE CIVIL DOCUMENTS
    qc20,        # OFFICIAL DOCUMENTS SHOULD INCLUDE THIRD GENDER OPTION
    qc15_1,      # LGB STATEMENTS - SAME RIGHTS AS HETEROSEXUAL
    qc15_2,      # LGB STATEMENTS - NOTHING WRONG WITH SAME SEX RELATIONSHIPS
    qc15_3,      # LGB STATEMENTS - ALLOW SAME SEX MARRIAGE THROUGHOUT EUROPE
    qc4_7,       # JOB CANDIDATE DISADVANTAGE: GENDER
    qc4_8,       # JOB CANDIDATE DISADVANTAGE: GENDER IDENTITY
    qc5_3,       # ACTIONS AGAINST DISCRIMINATION - JOINED ASSOCIATION/CAMPAIGN
    qc7,          # EFFORTS TO FIGHT DISCRIMINATION - RATING
    country_name
  )

#convert to factor, for further processing
legal_framework_data<- legal_framework_data %>%
  mutate(across(everything(), as.factor))

str(legal_framework_data)


```




```{r}
#| echo: false

qc19_summary <- legal_framework_data %>%
  group_by(country_name, qc19) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

```


Transgender Rights in Europe: Which Countries Support Legal Gender Changes?

```{r}
qc19_summary_sorted <- qc19_summary %>%
  filter(qc19 == "1") %>%
  arrange(percentage)

ggplot(qc19_summary_sorted, aes(x = reorder(country_name, percentage), y = percentage, fill = percentage)) +
  geom_bar(stat = "identity", color = "black", width = 0.7) +
  coord_flip() +
  scale_fill_gradient(low = "lightblue", high = "darkblue") + # Farbverlauf für bessere Lesbarkeit
  geom_text(aes(label = paste0(round(percentage, 1), "%")), hjust = -0.2, size = 5, fontface = "bold") +
  labs(
    title = "Support for Transgender Document Change (qc19) per Country",
    x = "Country",
    y = "Support Rate (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.position = "none"
  )
```


```{r}
qc19_20_summary <- legal_framework_data %>%
  filter(qc19 %in% c("1", "2"), qc20 %in% c("1", "2")) %>% # Nur Ja/Nein-Antworten
  group_by(country_name, qc19, qc20) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(qc19_20_summary, aes(x = reorder(country_name, -percentage), y = percentage, fill = qc20)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  labs(
    title = "Support for Transgender Document Change (qc19) vs. Third Gender Option (qc20)",
    x = "Country", y = "Percentage", fill = "Third Gender Option Support"
  ) +
  theme_minimal()
```




```{r}
lgbt_index <- read.csv("lgbt-rights-index.csv")
```



