---
title: "Group_Project"
format: html
editor: visual
---

## Quarto

## Libraries:

```{r}
library(haven)
library(dplyr)
library(ggplot2)
library(stringr) 
library(lubridate)
library(corrr)
library(ggplot2)
library(ggcorrplot)
library(tidyr)
library(scales)
```

Data:

```{r}
#| echo: false
data <- read_dta("ZA7575.dta")
```

```{r}
#| echo: false
summary(data)
head(data)


column_info <- data.frame(
  column_name = colnames(data),
  column_label = sapply(data, function(x) attr(x, "label"))
)

column_info

data <- data %>%
  mutate(country_name = case_when(
    q1_1  == 1 ~ "Belgium",
    q1_2  == 1 ~ "Denmark",
    q1_3  == 1 ~ "Germany",
    q1_4  == 1 ~ "Greece",
    q1_5  == 1 ~ "Spain",
    q1_6  == 1 ~ "France",
    q1_7  == 1 ~ "Ireland",
    q1_8  == 1 ~ "Italy",
    q1_9  == 1 ~ "Luxembourg",
    q1_10 == 1 ~ "Netherlands",
    q1_11 == 1 ~ "Portugal",
    q1_12 == 1 ~ "United Kingdom",
    q1_13 == 1 ~ "Austria",
    q1_14 == 1 ~ "Sweden",
    q1_15 == 1 ~ "Finland",
    q1_16 == 1 ~ "Cyprus",
    q1_17 == 1 ~ "Czech Republic",
    q1_18 == 1 ~ "Estonia",
    q1_19 == 1 ~ "Hungary",
    q1_20 == 1 ~ "Latvia",
    q1_21 == 1 ~ "Lithuania",
    q1_22 == 1 ~ "Malta",
    q1_23 == 1 ~ "Poland",
    q1_24 == 1 ~ "Slovakia",
    q1_25 == 1 ~ "Slovenia",
    q1_26 == 1 ~ "Bulgaria",
    q1_27 == 1 ~ "Romania",
    q1_28 == 1 ~ "Croatia",
    q1_29 == 1 ~ "Other countries",
    q1_30 == 1 ~ "DK",
    TRUE ~ NA_character_  
  ))




```

## legal framework

QC19:Do you think that transgender persons should be able to change their civil documents to match their inner gender identity?(1=yes, 2=no,3=dk) QC20: Do you believe that official documents, like passports and birth certificates, should have a third option, such as X or O (other) beside male (M) and female (F)for those persons who do not identify as female and male? (1=yes, 2=no,3=dk) QC15_1: to what extend to do you agree or disagree with the following statement? Gay, lesbian and bisexual people should have the same rights as heterosexual people (1 2 3 4 5) QC15_2: to what extend to do you agree or disagree with the following statement? here is nothing wrong in a sexual relationship between two persons of the same sex(1 2 3 4 5) QC15_3: to what extend to do you agree or disagree with the following statement? Same sex marriages should be allowed throughout europe (1 2 3 4 5) QC4_7: In (OUR COUNTRY) when a company wants to hire someone and has the choice between two candidates with equal skills and qualifications, which of the following criteria may, in your opinion, put one candidate at a disadvantage? The candidate’s gender or sex (man or woman) QC4_8: In (OUR COUNTRY) when a company wants to hire someone and has the choice between two candidates with equal skills and qualifications, which of the following criteria may, in your opinion, put one candidate at a disadvantage? The candidate’s gender identity (being transgender) QC5_3: in the last 12 month have you done: You have joined an association or campaign that defends people against discrimination? (1=yes, 2=no,3=dk)

```{r}
#| echo: false

legal_framework_data <- data %>%
  select(
    qc19,        # TRANSGENDER - CHANGE CIVIL DOCUMENTS
    qc20,        # OFFICIAL DOCUMENTS SHOULD INCLUDE THIRD GENDER OPTION
    qc15_1,      # LGB STATEMENTS - SAME RIGHTS AS HETEROSEXUAL
    qc15_2,      # LGB STATEMENTS - NOTHING WRONG WITH SAME SEX RELATIONSHIPS
    qc15_3,      # LGB STATEMENTS - ALLOW SAME SEX MARRIAGE THROUGHOUT EUROPE
    qc4_7,       # JOB CANDIDATE DISADVANTAGE: GENDER
    qc4_8,       # JOB CANDIDATE DISADVANTAGE: GENDER IDENTITY
    qc5_3,       # ACTIONS AGAINST DISCRIMINATION - JOINED ASSOCIATION/CAMPAIGN
    country_name
  )

#convert to factor, for further processing 
legal_framework_data<- legal_framework_data %>%
  mutate(across(everything(), as.factor))


str(legal_framework_data)


desc_stats <- legal_framework_data %>%
  summarise(
    # Binary (0/1: y/n, 3 = idk)
    Mean_Binary_qc19 = mean(as.numeric(qc19), na.rm = TRUE),  
    Mean_Binary_qc20 = mean(as.numeric(qc20), na.rm = TRUE),  
    
    # Likert-Skalen (1-5)
    Mean_Likert_qc15_1 = mean(as.numeric(qc15_1), na.rm = TRUE),  
    Mean_Likert_qc15_2 = mean(as.numeric(qc15_2), na.rm = TRUE),
    Mean_Likert_qc15_3 = mean(as.numeric(qc15_3), na.rm = TRUE)
  )

print(desc_stats)

```

The analysis shows clear differences in support for various aspects of LGBT+ rights. It is particularly noticeable that both transgender-relevant topics and LGB rights tend to be rated positively overall, but with certain differences in the intensity of approval. The average agreement with the possibility of transgender people being able to adapt their official documents to their gender identity (qc19) is 1.59. This indicates that a majority of respondents are in favor of this measure, although “Don't Know” responses could play a role. The introduction of a third gender option on official documents (qc20) received slightly less approval with an average value of 1.72, which indicates that this measure is discussed somewhat more controversially.

In comparison, the values for the general equality of LGB people (qc15_1,, qc15_2 , qc15_3) are more on the approving side of the scale with average values between 2.17 and 2.38 (1 = strongly approve, 5 = strongly disapprove). This means that the majority of respondents tend to agree with these statements. However, the values are not extremely low, which indicates that there is still a certain skepticism or reluctance in parts of the population.

Identifiying how much people choosed the third option: i dont know in percentage:

```{r}
dk_percentage <- legal_framework_data %>%
  summarise(
    qc19 = sum(qc19 == 3, na.rm = TRUE) / n() * 100,
    qc20 = sum(qc20 == 3, na.rm = TRUE) / n() * 100
  )

dk_percentage

```

qc19 (Transgender people should be able to adapt official documents) → 11.95% qc20 (Third gender option in official documents) → 11.62%

Country differences:

```{r}

dk_summary <- legal_framework_data %>%
  group_by(country_name) %>%
  summarise(
    DK_qc19 = sum(qc19 == 3, na.rm = TRUE) / n() * 100,
    DK_qc20 = sum(qc20 == 3, na.rm = TRUE) / n() * 100
  ) %>%
  pivot_longer(cols = c(DK_qc19, DK_qc20), names_to = "Question", values_to = "DK_Percentage")


ggplot(dk_summary, aes(x = reorder(country_name, DK_Percentage), y = DK_Percentage, fill = Question)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  coord_flip() +
  scale_fill_manual(values = c("DK_qc19" = "pink", "DK_qc20" = "orange"),
                    labels = c("qc19: Transgender documentchange", "qc20: third genderoption")) +
  labs(title = "Don't Know Rate ",
       x = "country", y = "% of dont know answers",
       fill = "Frage") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, face = "bold"))



```

-\> yyou can see clear differences in the response rates. While in bulgaria every 5. person selected "dont know" for transgender document change (qc19) in Belgium only less than 5% choosed this option.

```{r}
#| echo: false



```

Transgender Rights in Europe: Which Countries Support Legal Gender Changes?

```{r}
qc19_summary_sorted <- legal_framework_data %>%
  group_by(country_name) %>%
  mutate(total_responses = n()) %>%  # Gesamtanzahl der Antworten pro Land berechnen
  group_by(country_name, qc19) %>%
  summarise(count = n(), total_responses = first(total_responses), .groups = "drop") %>%
  mutate(percentage = (count / total_responses) * 100) %>%
  filter(qc19 == "1") %>%
  arrange(percentage)


ggplot(qc19_summary_sorted, aes(x = reorder(country_name, percentage), y = percentage, fill = percentage)) +
  geom_bar(stat = "identity", color = "black", width = 0.7) +
  coord_flip() +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), hjust = -0.2, size = 5, fontface = "bold") +
  labs(
    title = "Support for Transgender Document Change (qc19) per Country",
    x = "Country",
    y = "Support Rate (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.position = "none"
  )


ggplot(qc19_summary_sorted, aes(x = reorder(country_name, percentage), y = percentage, fill = percentage)) +
  geom_bar(stat = "identity", color = "black", width = 0.7) +
  coord_flip() +
  scale_fill_gradientn(
    colors = c("red", "orange", "yellow", "green", "blue", "purple"), 
    values = rescale(c(min(qc19_summary_sorted$percentage), max(qc19_summary_sorted$percentage))),
    guide = "colorbar"
  ) +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), hjust = -0.2, size = 5, fontface = "bold") +
  labs(
    title = "Support for Transgender Document Change (qc19) per Country",
    x = "Country",
    y = "Support Rate (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.position = "right"
  )


```

Results: The Netherlands (88%), Spain (76%) and Malta (76%) are the countries with the highest approval ratings.Countries such as Germany (62%), Sweden (62%) and Austria (57%) also show a majority in favor of changing the law, but at a lower level than the frontrunners. In these countries, there appears to be ongoing social debate and differing opinions. Bulgaria (12%), Hungary (14%), Romania (22%) and Slovakia (24%) have the lowest approval ratings.

Because there arent so many variables for legal framework, we included the lgtb-rights index-data. The Data is from: https://ourworldindata.org/grapher/lgbt-rights-index?time=2012&region=Europe The index captures to which extent LGBT+ people have the same rights as straight and cisgender people. It combines 18 individual policies, such as the legality of same-sex relationships, marriage, and gender marker changes. Higher values indicate more rights, negative values regressive policies progressive policies receiving a positive score and regressive policies receiving a negative score: -5 to +13. Among the progressive policies are: Same-sex sexual acts Legal, Equal Age of Consent, Employment Discrimination, Hate Crime Protections, Incitement to Hatred, Civil Unions, Marriage Equality, oint Adoptions, Gender Marker Change, LGB Military, Transgender Military, Ban on Conversion Therapies, and Ban on Gender Assignment Surgeries on Children

```{r}
lgbt_index <- read.csv("lgbt-rights-index.csv")
str(lgbt_index)

lgbt_index <- lgbt_index |> filter(Year== 2019)
```

```{r}

merged_data <- legal_framework_data %>%
  inner_join(lgbt_index, by = c("country_name" = "Entity"))

ggplot(merged_data, aes(x = LGBT..Policy.Index, y = reorder(country_name, LGBT..Policy.Index), color = LGBT..Policy.Index)) +
  geom_point(size = 4) +
  scale_color_gradient(low = "red", high = "blue") +  
  labs(title = "LGBT Policy Index by Country",
       x = "LGBT Policy Index", y = "Country") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, face = "bold"))

```

Malta has the highest score, well above all other countries. Scandinavian & Western European countries (Sweden, Netherlands, UK, Germany, France) have relatively high values. Eastern European countries (Latvia, Poland, Bulgaria, Romania, Hungary) have the lowest values in the LGBT Index.

We wanted to investigate whether there are significant differences in the laws in different parts of Europe. We therefore divided the countries into Eastern, Western, Southern and Northern Europe and compared the LGTB score.

```{r}
merged_data <- merged_data %>%
  mutate(region = case_when(
    country_name %in% c("Denmark", "Sweden", "Finland", "Estonia", "Latvia", "Lithuania") ~ "Northern Europe",
    country_name %in% c("Germany", "France", "Netherlands", "Belgium", "Luxembourg", 
                        "United Kingdom", "United Kingdom of Great Britain and Northern Ireland",  
                        "Ireland", "Austria") ~ "Western Europe",
    country_name %in% c("Spain", "Italy", "Portugal", "Greece", "Malta", "Cyprus") ~ "Southern Europe",
    country_name %in% c("Poland", "Hungary", "Czech Republic", "Slovakia", "Romania", 
                        "Bulgaria", "Croatia", "Slovenia") ~ "Eastern Europe",
    TRUE ~ "Other"
  ))


ggplot(merged_data, aes(x = region, y = LGBT..Policy.Index, fill = region)) +
  geom_boxplot() +
  labs(title = "LGBT Policy Index by Region",
       x = "Region", y = "LGBT Policy Index") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


             
```

Western Europe has the highest LGBT index, with a high median and a relatively small range. Northern Europe shows a large variance, with some countries having very high values and others in the middle. Southern Europe is lower overall than Northern and Western Europe, with a wide range of values. Eastern Europe has the lowest values, with a median of around 5 and lower protection rights for LGBT people.

```{r}
anova_result <- aov(LGBT..Policy.Index ~ region, data = merged_data)
summary(anova_result)
```

The ANOVA analysis shows a highly significant difference between the regions (p \< 2e-16), which means that the differences in the LGBT index between the regions are statistically significant.

Interpretation: LGBT policy differs systematically between the regions, with Western Europe and Northern Europe having significantly higher values than Eastern Europe and Southern Europe.

```{r}
cor_data <- merged_data %>%
  select(qc19, qc20, qc15_1, qc15_2, qc15_3,LGBT..Policy.Index) %>%
  mutate(across(everything(), as.numeric))  

cor_matrix <- cor(cor_data, use = "pairwise.complete.obs")

print(cor_matrix)

ggcorrplot(cor_matrix, method = "circle", type = "lower",
           lab = TRUE, lab_size = 4, colors = c("red", "white", "blue"),
           title = "Correlation Matrix of Selected Variables")
             
```

The correlation table highlights interesting connections between societal views on LGBTI+ rights and the LGBT Policy Index. The strongest positive correlation appears between these variables: qc15_1 which gauges equal rights for LGB people; qc15_2 which measures acceptance of same-sex relationships; and qc15_3 which covers marriage equality. These values range from 0.75 to 0.82. It indicates that nations with high acceptance of LGB rights generally back marriage equality too. There's a moderate correlation between qc19 which covers support for transgender document changes and qc20 which deals with a third gender option, clocking in at 0.59. This suggests that countries with advanced laws for trans people often favor a third gender option on official documents as well. But the correlations between qc19/qc20 and general LGBTI+ attitudes ‒ qc15\_\* ‒ are moderate (0.33–0.43). A positive attitude towards LGB rights often occurs alongside supportive views on trans issues.

```{r}

lm_model <- lm(LGBT..Policy.Index ~ qc19 + qc20 + qc15_1 + qc15_2 + qc15_3, data = merged_data)

summary(lm_model)
```

The results show that approval of LGBT rights is strongly associated with a higher LGBT Policy Index. the strongest negative effects come from rejection of “marriage for all” (-1.76 points) and same-sex relationships (-0.94 points). neutral answers (“don't know”) are often as negative as rejection - uncertainty is therefore a strong signal for less legal protection of LGBT+ rights. the transgender-specific questions (qc19, qc20) are also significant, but with somewhat weaker effects than LGB rights. the model is significant, but only explains approx. 26.5% of the variance - other factors also play a role.

### Cult and History

```{r}
library(httr)
library(dplyr)
library(rvest)
library(lubridate)
library(tidyr)
library(ggplot2)
```

to see the correlation between two numeric variables: notice that they are from two different datasets. it’s adjustable

cor(pca_data\$Attendance, pca_data_target\$Home_goal_diff, use = "complete.obs")

**Cultural Norms**

```{r}

```

### History

#### Accession Date

The accession date to the European Union might be used as a proxy for democratic practices and culture. To assess this, we will use our first external data for this part. Let's get the accession dates for all European Union member countries.

```{r}

eu_accession <- "https://en.wikipedia.org/wiki/Member_state_of_the_European_Union"
eu_accession_table <- 
  eu_accession %>% 
  read_html() %>% 
  html_elements("table") %>% 
  html_table(fill = TRUE) %>% 
  .[[2]]

# Clean column names
colnames(eu_accession_table) <- make.names(colnames(eu_accession_table))

# Preview
head(eu_accession_table)
```

Unfortunately, dates are characters and they contain written versions of the months. To fix this:

```{r}
# Clean and convert 'Accession' column to Date format
eu_accession_table <- eu_accession_table %>%
  mutate(Accession = gsub("\\s*Founder.*", "", Accession),  # Remove extra text
         Accession = dmy(Accession))  # Convert to Date format (Day-Month-Year)

# to make sure they are dates now
str(eu_accession_table$Accession)

# Let's put the foundation date of the EU for founding countries
eu_accession_table$Accession[is.na(eu_accession_table$Accession)] <- as.Date("1957-03-25")

# Let's get rid of the silly brackets next to country names too

eu_accession_table$Country <- gsub("\\[.*?\\]", "", eu_accession_table$Country)

print(eu_accession_table$Country)

eu_accession_table <-
  eu_accession_table |> 
  select(Country, ISO, Accession) |> 
  rename(country_name = Country)

save(eu_accession_table, file = "eu_accession_table.rda")
```

#### Form of governments

For descriptive analysis, we want to check if there is a pattern for the support of LGBT rights on the basis of government form.

```{r}
eu_gov_styles <- "https://en.wikipedia.org/wiki/List_of_European_Union_member_states_by_political_system"
eu_gov_styles_table <- 
  eu_gov_styles %>% 
  read_html() %>% 
  html_elements("table") %>% 
  html_table(fill = TRUE) %>% 
  .[[2]]

# keeping only the necessary columns 

eu_gov_styles_table <-
  eu_gov_styles_table |> 
  select(State, Government, 'Monarchy/Republic') |> 
  rename(country_name = State) |> 
  mutate(country_name = ifelse(country_name == "Czech Republic", "Czechia", country_name))

save(eu_gov_styles_table, file = "eu_gov_styles_table.rda")
```

#### Former socialist - Western bloc

Even more interestingly, we want to check how does the level of LGBT support vary across former socialist states and western bloc states.

This will be done manually because we were not able to find a single table indicating historical ideology data across all European countries. We were able to find the list of former socialist EU countries however the table was way too untidy. We then decided to take advantage of low number of countries and do it manually so that we will not lose this information.

The case of Germany is tricky, so we decided to evaluate it as a unique occurrence.

For the former socialist states, see: <https://en.wikipedia.org/wiki/List_of_socialist_states>

Bulgaria, Czechoslovakia (Now Czechia and Slovakia), Hungary, Poland, Romania, some former Soviet Union (for our case Lithuania, Latvia, and Estonia) states, and some former Yugoslavia (Slovenia, Croatia) states are the ones that used to be ruled under socialism and then joined the EU.

```{r}
# socialist bloc - now EU countries
socialist_bloc <- c("Bulgaria", "Croatia", "Czechia", "Estonia", "Hungary", 
                    "Latvia", "Lithuania", "Poland", "Romania", "Slovakia", "Slovenia")

# western bloc - now EU countries
western_bloc <- c("Austria", "Belgium", "Cyprus", "Denmark", "Finland", 
                  "France", "Greece", "Ireland", "Italy", "Luxembourg", 
                  "Malta", "Netherlands", "Portugal", "Spain", "Sweden", "United Kingdom") # adding the UK because it was a member at the time of survey



mixed_bloc <- c("Germany") # putting Germany into mixed for obvious reasons

countries <- c(socialist_bloc, western_bloc, mixed_bloc)
blocs <- c(rep("Socialist", length(socialist_bloc)), rep("Western", length(western_bloc)), rep("Mixed", length(mixed_bloc)))

eu_blocs <- data.frame(Country = countries, Bloc = blocs)
print(eu_blocs)

eu_blocs$Bloc_binary <- ifelse(eu_blocs$Bloc == "Socialist", 0, 1)

# to match the country_name column for joining tables later

eu_blocs <-
  eu_blocs |> 
  rename(country_name = Country)

# to match the name of Czechia for all tables

data <- data |> 
  mutate(country_name = ifelse(country_name == "Czech Republic", "Czechia", country_name))

save(data, file = "data.rda")
save(eu_blocs, file = "eu_blocs.rda")
```

#### Variables and some adjusting

```{r}

load("data.rda")
load("eu_blocs.rda")
load("eu_gov_styles_table.rda")
load("eu_accession_table.rda")

culture_hist_data <- data |> 
  select(
    country_name,
    d11r3,              # Age recoded with 7 categories
    sd3,                # Religious affiliation
    gen3,               # Gen before 1946
    gen4,               # Gen 1946-1964
    gen5,               # Gen 1965-1980
    gen6,               # Gen after 1980
    d10,                # Gender of the respondent
    qc19,               # The main question of the assignment
    sd1_7,              # Personal contact with a transgender person irl
    qc2t_3,             # experienced discriminiation irl, summed for all LGBT
    qc4t_3,             # job candidate disadvantage, summed for all LGBT 
    d1r2,               # political ideology of the respondent, cat 5
    qc6_2,              # would you like an elected politician to be homosexual
    qc6_2r,             # would you like an elected politician to be homosexual, rec
    qc6_10,             # would you like an elected politician to be transgender
    qc6_10r             # would you like an elected politician to be transgender, rec
  )

```

Let's factor the country name variable for future use

```{r}

culture_hist_data$country_name <- factor(culture_hist_data$country_name)
```

Throughout this part, we will try and make sense of attitudes towards LGBT community by making distinctions on the basis of political & historical characteristics of EU member countries. First, let's join our three external tables for this part.

-   European member countries by their allegiance to socialist or western bloc in the past
-   European member countries by the date they joined the Union
-   European member countries by the forms of government

```{r}
merged_cult_hist <- culture_hist_data |> 
  left_join(eu_blocs, by = "country_name") |> 
  left_join(eu_gov_styles_table, by = "country_name") |> 
  left_join(eu_accession_table, by = "country_name")

save(merged_cult_hist, file = "merged_cult_hist.rda")
```

Before actually proceeding to visualisations and explanatory analysis, let's see if we have any NA values across our table.

```{r}
colSums(is.na(merged_cult_hist))

```

Apparently we do have missing values. After careful investigation, we realised that one country is causing missing values after merging with external tables.

United Kingdom: Due to Brexit. The external data that shows information for the EU member states are updated and excluded the UK.

After a little manual adjustment we will be set to go.

Consequent to following below code, we end up having a complete dataset.

```{r}

# This will fill in the Accession and government form columns for the UK

merged_cult_hist <- merged_cult_hist |> 
  mutate(
    Accession = case_when(
      country_name == "United Kingdom" ~ as.Date("1973-01-01"),
      TRUE ~ Accession
    ),
    `Monarchy/Republic` = case_when(
      country_name == "United Kingdom" ~ "constitutional monarchy",
      TRUE ~ `Monarchy/Republic`
    )
  )

merged_cult_hist <- merged_cult_hist |> 
  select(-Government, -ISO)
save(merged_cult_hist, file = "merged_cult_hist.rda")
```

#### Gender

```{r}
# first, I want to recode gender variable for best practice
# the documentation of the questionnaire tells us that 1 = Man, 2 = Woman
# to follow the best practices, we will do woman = 0 and man = 1

merged_cult_hist <- merged_cult_hist |> 
  mutate(
    d10 = case_when(
      as.character(d10) == "1" ~ "1",
      as.character(d10) == "2" ~ "0",
      TRUE ~ as.character(d10)
    ),
    d10 = as.numeric(d10)
  )
```

Now let's plot the relationship between gender and support for LGBT documentation rights.

Yielded results indicate that women are slightly more supportive of documentation rights of LGBT communities than men. Rejecting those rights is more common amongst men compared to women, and the response "Don't know" is more frequently seen amongst women than men.

```{r}

summary_df <- merged_cult_hist |> 
  group_by(d10, qc19) |> 
  summarise(Count = n()) |> 
  group_by(d10) |> 
  mutate(Percentage = Count / sum(Count) * 100) |> 
  ungroup() |> 
  mutate(qc19 = factor(qc19, levels = c(1, 2, 3), labels = c("Yes", "No", "Don't know")),
         d10 = factor(d10, levels = c(1, 0), labels = c("Men", "Women")))

# visualizing

ggplot(summary_df, aes(x = qc19, y = Percentage, fill = d10)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = sprintf("%.1f%%", Percentage)), 
            position = position_dodge(width = 0.9), vjust = -0.5) +
  scale_fill_manual(values = c("Men" = "#a0d6b3", "Women" = "#e6886f")) +
  labs(title = "Support for LGBT Documentation Rights",
       x = "", y = "Percentage", fill = "") +
  theme_minimal() +
  theme(plot.title = element_text(size = 12, hjust = 0.5))
```

#### Generations

Another useful insight that we can draw from the survey data is the levels of generational support and rejection of LGBT rights relating to identification issues. The barplot below expectedly reveals that the greatest support for the issue comes from the generation that was born after 1980. Gradually decreasing backwards, the lowest support level is posted by before 1946 generation.

Rejection, on the other hand, shows a quite homogenous distribution amongst all generations. However, the greatest rejection is demonstrated by the generation before 1946, whereas the lowest rejection percentage belongs to after-1980 generation.

With around 20 per cent, before-1946 generation expressed with the highest percentage that they do not know about the topic, which could be viewed as an expected outcome.

```{r}

# before plotting, let's summarize each generation and create percentages for support/rejection levels

gen_summary <- rbind(
  merged_cult_hist |> 
    filter(gen3 == 1) |> 
    group_by(qc19) |> 
    summarise(Count = n()) |> 
    mutate(Generation = "Before 1946", Percentage = Count / sum(Count) * 100),
  merged_cult_hist |> 
    filter(gen4 == 1) |> 
    group_by(qc19) |> 
    summarise(Count = n()) |> 
    mutate(Generation = "1946-1964", Percentage = Count / sum(Count) * 100),
  merged_cult_hist |> 
    filter(gen5 == 1) |> 
    group_by(qc19) |> 
    summarise(Count = n()) |> 
    mutate(Generation = "1965-1980", Percentage = Count / sum(Count) * 100),
  merged_cult_hist |> 
    filter(gen6 == 1) |> 
    group_by(qc19) |> 
    summarise(Count = n()) |> 
    mutate(Generation = "After 1980", Percentage = Count / sum(Count) * 100)
) |> 
  mutate(qc19 = factor(qc19, levels = c(1, 2, 3), labels = c("Yes", "No", "DK")),
         Generation = factor(Generation, levels = c("Before 1946", "1946-1964", "1965-1980", "After 1980")))

# visualizing

ggplot(gen_summary, aes(x = qc19, y = Percentage, fill = Generation)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = sprintf("%.1f%%", Percentage)), 
            position = position_dodge(width = 0.9), vjust = -1, size = 3) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "LGBT Documentation Rights - Generation", x = "", y = "Support percentage", fill = "Generation") +
  theme_minimal()

```

#### Political ideology - individual

##### Support for LGBT documents to be changed - Political ideology:

Self political report gives us hints about the distribution of LGBT support across different groups in the society, therefore signaling the political culture within the EU. According to the below plot, most supportive groups are those who identify themselves to be far and moderately left, with the moderate left leading. These groups are tightly followed by centrists, whereas the least support and strongest rejection are seen in the far right groups. Amongst those who do not want to reveal their political affiliation, support level is similar to far right. This group could be interesting for making further investigation and comments.

```{r}
# Summarizing the data to plot 
summary_df <- merged_cult_hist |> 
  group_by(d1r2, qc19) |> 
  summarise(Count = n()) |> 
  group_by(d1r2) |> 
  mutate(Percentage = Count / sum(Count) * 100) |> 
  ungroup() |> 
  mutate(d1r2 = factor(d1r2, levels = c(1, 2, 3, 4, 5, 9),
                       labels = c("(1-2) Left", "(3-4)", "(5-6) Centre", "(7-8)", "(9-10) Right", "DK/Refusal")),
         qc19 = factor(qc19, levels = c(1, 2, 3), labels = c("Yes", "No", "DK")))


# let's create a stacked bar plot

ggplot(summary_df, aes(x = d1r2, y = Percentage, fill = qc19)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Yes" = "#4B87F0", "No" = "lightblue", "DK" = "#FF6B6B")) +
  labs(title = "QC19 Responses by Political Placement",
       x = "Political Placement", y = "Percentage", fill = "Response") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

##### Support for LGBT documents to be changed - Political ideology (by genders)

The following plot shows the same results but by gender. Results do not show any different implications than the previous one.

```{r}
# first, summarizing the data to plot
summary_df <- merged_cult_hist |> 
  group_by(d1r2, qc19, d10) |> 
  summarise(Count = n()) |> 
  group_by(d1r2, d10) |> 
  mutate(Percentage = Count / sum(Count) * 100) |> 
  ungroup() |> 
  mutate(d1r2 = factor(d1r2, levels = c(1, 2, 3, 4, 5, 9),
                       labels = c("(1-2) Left", "(3-4)", "(5-6) Centre", "(7-8)", "(9-10) Right", "DK/Refusal")),
         qc19 = factor(qc19, levels = c(1, 2, 3), labels = c("Yes", "No", "DK")),
         d10 = factor(d10, levels = c(1, 0), labels = c("Men", "Women")))

# let's create a stacked bar plot

ggplot(summary_df, aes(x = d1r2, y = Percentage, fill = qc19)) +
  geom_bar(stat = "identity") +  
  facet_wrap(~d10, nrow = 1) +
  scale_fill_manual(values = c("Yes" = "lightblue", "No" = "#D6A8E6", "DK" = "#B0BEC5")) +
  labs(title = "QC19 Responses by Self Political Placement, Split by Gender",
       x = "Political Placement", y = "Percentage", fill = "Response") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#### Religiousity

Religious affilition could be considered to shape people's opinions on LGBT communities, as religions often have strong opinions regarding those issues. That is why we wanted to see how support for LGBT ID change varies across different religious affiliations.

Orthodox Christians and Muslims are immediately caught by the eye in that they post the lowest proportion of support for the question at hand. They are closely followed by Catholic Christians and Hindus.

Non-believers, Atheists, and Protestans show considerable support.

Sikhs, finally, is the group that is most supportive of LGBT documentation rights.

Findings from this plot are in line with the findings of historical data where we plotted past affiliation with socialism. Majority of previously-socialist countries are known to have a big proportion of Orthodox population.

```{r}

# Recoding all Muslim categories since. They are a minority and summing them up could be better

summary_df <- merged_cult_hist |> 
  mutate(sd3_recode = case_when(
    sd3 %in% c(6, 7, 8) ~ 7,  #  sunni, shia, and other into one muslim level
    TRUE ~ sd3
  )) %>%
  group_by(sd3_recode, qc19) |> 
  summarise(Count = n()) |> 
  group_by(sd3_recode) |> 
  mutate(Percentage = Count / sum(Count) * 100) |> 
  ungroup() |> 
  mutate(sd3_recode = factor(sd3_recode, levels = c(1, 2, 3, 4, 5, 7, 9, 10, 11, 12, 13, 14, 15, 16),
                             labels = c("Catholic", "Orthodox Christian", "Protestant", "Other Christian", "Jewish", "Muslim", "Sikh", "Buddhist", "Hindu", "Atheist", "Non-believer/Agnostic", "Other", "Refusal", "DK")),
         qc19 = factor(qc19, levels = c(1, 2, 3), labels = c("Yes", "No", "DK")))

# plotting
ggplot(summary_df, aes(x = sd3_recode, y = Percentage, fill = qc19)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Yes" = "#a0d6b3", "No" = "#e6886f", "DK" = "#B0BEC5")) +
  labs(title = "QC19 Responses by Religion/Beliefs",
       x = "Religion/Beliefs", y = "Percentage", fill = "Response") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#### Political history - Socialist vs Western

In this part, we will visualize the support for LGBT people to be able to change their IDs by political history of all EU member states. The 'socialist' vs 'western' or 'democratic' distinction has its roots in the Cold War era which made its mark to the 20th century.

Results show that there is a clear distinction between countries from each legacy. The socialist bloc demonstrates a rejection of such right for LGBT people with slight majority. However, there is a considerable support as well.

Western-affiliated countries, on the other hand, are highly supportive of LGBT documentation rights with a clear majority. With a rejection proportion around one-third of approval, there is a considerable rejection as well.

'Mixed' label represents Germany due to the division of the country by socialist and western administrations. We wanted to visualize it separately for obvious reasons. In Germany, support for LGBT ID change overrides other two options with a decisive majority. This is expected considering the commitment of the country to democratic values and practices through the years.

```{r}

summary_df <- as.data.frame(table(merged_cult_hist$qc19, merged_cult_hist$Bloc))
colnames(summary_df) <- c("qc19", "Bloc", "Count")

# mapping the target variable to labels
summary_df$qc19_label <- factor(summary_df$qc19, levels = c(1, 2, 3), labels = c("Yes", "No", "DK"))

# plotting

ggplot(summary_df, aes(x = Bloc, y = Count, fill = qc19_label)) +
  geom_bar(stat = "identity", color = "black", size = 0.5) +
  scale_fill_manual(values = c("Yes" = "#a1d99b", "No" = "#F44336", "DK" = "#B0BEC5")) +
  labs(title = "Transgender Document Change Support-Historical Bloc", x = "Bloc", y = "Count", fill = "Response") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        legend.title.position = "top",
        legend.position = "top",
        legend.key.size = unit(0.3, "cm"),
        panel.grid.major.x = element_blank())

# to see the precise numbers

merged_cult_hist |> 
  group_by(Bloc, qc19) |> 
  summarise(Count = n()) |> 
  mutate(qc19 = factor(qc19, levels = c(1, 2, 3), labels = c("Yes", "No", "DK"))) |> 
  print()
```

## legal framework

QC19:Do you think that transgender persons should be able to change their civil documents to match their inner gender identity?(1=yes, 2=no,3=dk) QC20: Do you believe that official documents, like passports and birth certificates, should have a third option, such as X or O (other) beside male (M) and female (F)for those persons who do not identify as female and male? (1=yes, 2=no,3=dk) QC15_1: to what extend to do you agree or disagree with the following statement? Gay, lesbian and bisexual people should have the same rights as heterosexual people (1 2 3 4 5) QC15_2: to what extend to do you agree or disagree with the following statement? here is nothing wrong in a sexual relationship between two persons of the same sex(1 2 3 4 5) QC15_3: to what extend to do you agree or disagree with the following statement? Same sex marriages should be allowed throughout europe (1 2 3 4 5) QC4_7: In (OUR COUNTRY) when a company wants to hire someone and has the choice between two candidates with equal skills and qualifications, which of the following criteria may, in your opinion, put one candidate at a disadvantage? The candidate’s gender or sex (man or woman) QC4_8: In (OUR COUNTRY) when a company wants to hire someone and has the choice between two candidates with equal skills and qualifications, which of the following criteria may, in your opinion, put one candidate at a disadvantage? The candidate’s gender identity (being transgender) QC5_3: in the last 12 month have you done: You have joined an association or campaign that defends people against discrimination? (1=yes, 2=no,3=dk) QC7: Using a scale from 1 to 10, please tell me if you think that the efforts made in (OUR COUNTRY) to fight all forms of discrimination are effective. ‘1’ means you consider that these efforts are "not at all effective", and ‘10’ that these efforts are "very effective".

```{r}
#| echo: false

legal_framework_data <- data %>%
  select(
    qc19,        # TRANSGENDER - CHANGE CIVIL DOCUMENTS
    qc20,        # OFFICIAL DOCUMENTS SHOULD INCLUDE THIRD GENDER OPTION
    qc15_1,      # LGB STATEMENTS - SAME RIGHTS AS HETEROSEXUAL
    qc15_2,      # LGB STATEMENTS - NOTHING WRONG WITH SAME SEX RELATIONSHIPS
    qc15_3,      # LGB STATEMENTS - ALLOW SAME SEX MARRIAGE THROUGHOUT EUROPE
    qc4_7,       # JOB CANDIDATE DISADVANTAGE: GENDER
    qc4_8,       # JOB CANDIDATE DISADVANTAGE: GENDER IDENTITY
    qc5_3,       # ACTIONS AGAINST DISCRIMINATION - JOINED ASSOCIATION/CAMPAIGN
    qc7,          # EFFORTS TO FIGHT DISCRIMINATION - RATING
    country_name
  )


legal_framework_data<- legal_framework_data %>%
  mutate(across(everything(), as.factor))

str(legal_framework_data)


```

```{r}
#| echo: false

qc19_summary <- legal_framework_data %>%
  group_by(country_name, qc19) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(qc19_summary %>% filter(qc19 == "1"), aes(x = reorder(country_name, -percentage), y = percentage)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Support for transgender document change (qc19) per country",
       x = "Country", y = "Rate (%)") +
  theme_minimal()

```
